---
title: "Project 2"
author: "Javeria Fiaz"
date: "2023-06-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Delete Environment
```{r clear environment}
rm(list = ls())
rm()
```

## R Libraries
Load libraries

```{r libraries}
library(readr)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(lubridate)
```

## R CSV Files
Upload CSV Files

```{r upload csv data}
df_order_item <- read_csv("F:/JAVERIA FIAZ/June 2023/Portfolio Project 2/R code/Maven_csv/order_item.csv")
df_order_item_refund <- read_csv("F:/JAVERIA FIAZ/June 2023/Portfolio Project 2/R code/Maven_csv/order_item_refund.csv")
df_orders <- read_csv("F:/JAVERIA FIAZ/June 2023/Portfolio Project 2/R code/Maven_csv/orders.csv")
df_products <- read_csv("F:/JAVERIA FIAZ/June 2023/Portfolio Project 2/R code/Maven_csv/products.csv")
df_website_sessions <- read_csv("F:/JAVERIA FIAZ/June 2023/Portfolio Project 2/R code/Maven_csv/website _sessions.csv")
df_website_pageviews <- read_csv("F:/JAVERIA FIAZ/June 2023/Portfolio Project 2/R code/Maven_csv/website_pageviews.csv")
```


## Sample: save as cSV file
```{r}
write.csv(website_session_t1, file = "I:/DATA ANALYST/Portfolio Project 2/csv/website_session_t1.csv", row.names = FALSE)
```


## R Code
Total Sessions and users by utm_source and utm_campaign
```{r utm_url}
df_website_sessions |> 
group_by(utm_source,utm_campaign) |> 
summarise(total_sessions = n_distinct(website_session_id),
          total_users = n_distinct(user_id)) |> 
arrange(desc(total_sessions)) |>
  view()-> utm_url

```

New and Repeated Customers
```{r}
# Number of sessions for each user
df_website_sessions |> 
  group_by(user_id)  |> 
  summarize(total_sessions = n()) -> user_sessions 

user_sessions  |> 
  filter(total_sessions == 1)  |> 
  summarise(count = n()) -> new_customers

user_sessions  |> 
  filter(total_sessions > 1)  |> 
  summarise(count = n()) -> repeat_customers

View(new_customers)
View(repeat_customers)

```

## UTM_Source

Total users and total sessions by each source
```{r}
df_website_sessions |> 
  group_by(utm_source) |> 
  summarise(total_sessions = n_distinct(website_session_id),
            total_user = n_distinct(user_id)) |>
 view()-> utm_source

write.csv(utm_source , file = 
            "F:/JAVERIA FIAZ/June 2023/Portfolio Project 2/csv/utm_source.csv", row.names = FALSE)

```

Most used device_type by each source
````{r}
# Most used device_type by each source
df_website_sessions |> 
  group_by(utm_source, device_type) |>
  summarise(
   device_used = table(device_type)) -> s_device 
  view()
  
write.csv(s_device  , file = 
            "F:/JAVERIA FIAZ/June 2023/Portfolio Project 2/csv/s_device.csv", row.names = FALSE)
  
```

New and Repeated customer by each source
```{r}
# Add a column for yaer
#df_website_sessions$years <- format(df_website_sessions$created_at, "%Y")

# Number of sessions for each user
s_user_sessions <- df_website_sessions |> 
  group_by(user_id)  |> 
  summarize(total_sessions = n())  |> 
  left_join(
    df_website_sessions |> 
      group_by(user_id) |> 
      summarize(utm_sources = paste(unique(utm_source), collapse = ", ")),
    by = "user_id"
  )

# New customers (users with only one session)
s_new_customers <- s_user_sessions |> 
  filter(utm_sources %in% c("NULL", "gsearch", "bsearch", "socialbook")) |> 
  group_by(utm_sources) |> 
  summarize(new_users = sum(total_sessions == 1)) 

write.csv(s_new_customers , file = 
            "F:/JAVERIA FIAZ/June 2023/Portfolio Project 2/csv/s_new_customers.csv", row.names = FALSE)


# Repeat customers (users with more than one session)
s_repeat_customers <- s_user_sessions |> 
  filter(utm_sources %in% c("NULL", "gsearch", "bsearch", "socialbook")) |> 
  group_by(utm_sources) |> 
  summarize(repeat_users = sum(total_sessions > 1))

write.csv(s_repeat_customers , file = "F:/JAVERIA FIAZ/June 2023/Portfolio Project 2/csv/s_repeat_customers.csv", row.names = FALSE)

```


Traffic to order conversion for utm_source
```{r}
# Conversion Rate = (Total number of conversions / Total number of sessions)
# For example, if you had 50 conversions from 1,000 interactions, your conversion rate would be 5%, since 50 รท 1,000 = 5%.
# Higher CVR indicates that a higher percentage of website visitors or users are taking the desired action (such as making a purchase, signing up, or submitting a form)

df_website_sessions  |> 
  left_join(df_orders, by = "website_session_id")  |> 
  group_by(utm_source)  |> 
  summarize(
    sessions = n_distinct(website_session_id),
    orders = n_distinct(order_id),
    session_order_CVR = round((n_distinct(order_id) / n_distinct(website_session_id)*100), 1)
  )  |> 
  arrange(desc(session_order_CVR)) -> s_traffic_order_cvr 

write.csv(s_traffic_order_cvr , file = "F:/JAVERIA FIAZ/June 2023/Portfolio Project 2/csv/s_traffic_order_cvr.csv", row.names = FALSE)

```

Bounce Rate for utm_source
```{r}
df_website_pageviews %>%
  inner_join(df_website_sessions, by = "website_session_id") %>%
  group_by(utm_source, website_session_id) %>%
  summarise(pageviews = n_distinct(pageview_url)) %>%
  group_by(utm_source) %>%
  summarise(single_page = sum(pageviews == 1),
            bounce_rate = single_page / n() * 100) -> s_bounce_rate
  View()
  
  write.csv(s_bounce_rate, file = "F:/JAVERIA FIAZ/June 2023/Portfolio Project 2/csv/s_bounce_rate.csv", row.names = FALSE)

```

Show Traffic by each source
```{r for haet map display}
df_website_sessions  |> 
  mutate(hour = hour(created_at),
         hour_12 = format(created_at, "%I %p")) |> 
  group_by(utm_source, hour_12)  |> 
  summarize(count = n()) -> s_traffic_time
  
  
write.csv(s_traffic_time, file = "F:/JAVERIA FIAZ/June 2023/Portfolio Project 2/csv/s_traffic_time.csv", row.names = FALSE)

```

## UTM_Campaign

Total users and total sessions by each campaign
```{r}
df_website_sessions |> 
  group_by(utm_campaign) |> 
  summarise(total_sessions = n_distinct(website_session_id),
            total_user = n_distinct(user_id)) -> utm_campaign
 view()

write.csv(utm_campaign, file = 
            "F:/JAVERIA FIAZ/June 2023/Portfolio Project 2/csv/utm_campaign.csv", row.names = FALSE)

```

Most used device_type by each campaign
````{r}
# Most used device_type by each campaign
df_website_sessions |> 
  group_by(utm_campaign, device_type) |>
  summarise(
   device_used = table(device_type)) -> c_device 
  view()
  
write.csv(c_device  , file = 
            "F:/JAVERIA FIAZ/June 2023/Portfolio Project 2/csv/c_device.csv", row.names = FALSE)
  
```

New and Repeated customer by each campaign
```{r}

# Number of sessions for each user
df_website_sessions |> 
  group_by(user_id)  |> 
  summarize(total_sessions = n())  |> 
  left_join(
    df_website_sessions |> 
      group_by(user_id) |> 
      summarize(utm_campaigns = paste(unique(utm_campaign), collapse = ", ")),
    by = "user_id"
  ) -> c_user_sessions 

# New customers (users with only one campaign)
c_user_sessions |> 
  filter(utm_campaigns %in% c("nonbrand", "brand", "NULL", "desktop_targeted", "pilot")) |>
  group_by(utm_campaigns) |> 
  summarize(new_users = sum(total_sessions == 1)) -> c_new_customers

write.csv(c_new_customers, file = "F:/JAVERIA FIAZ/June 2023/Portfolio Project 2/csv/c_new_customers.csv", row.names = FALSE)


# Repeat customers (users with more than one campaign)
c_user_sessions |> 
  filter(utm_campaigns %in% c("nonbrand", "brand", "NULL", "desktop_targeted", "pilot")) |>
  group_by(utm_campaigns) |> 
  summarize(repeat_users = sum(total_sessions > 1)) -> c_repeat_customers

write.csv(c_repeat_customers, file = "F:/JAVERIA FIAZ/June 2023/Portfolio Project 2/csv/c_repeat_customers.csv", row.names = FALSE)

```

Traffic to order conversion for utm_campaign
```{r}
c_traffic_order_cvr <- df_website_sessions %>%
  left_join(df_orders, by = "website_session_id") %>%
  group_by(utm_campaign) %>%
  summarize(
    sessions = n_distinct(website_session_id),
    orders = n_distinct(order_id),
    session_order_CVR = round((n_distinct(order_id) / n_distinct(website_session_id)*100), 1)
  ) %>%
  arrange(desc(session_order_CVR))

print(c_traffic_order_cvr)

write.csv(c_traffic_order_cvr, file = "F:/JAVERIA FIAZ/June 2023/Portfolio Project 2/csv/c_traffic_order_cvr.csv", row.names = FALSE)

```

Bounce Rate for utm_campaign
```{r}
df_website_pageviews %>%
  inner_join(df_website_sessions, by = "website_session_id") %>%
  group_by(utm_campaign, website_session_id) %>%
  summarise(pageviews = n_distinct(pageview_url)) %>%
  group_by(utm_campaign) %>%
  summarise(single_page = sum(pageviews == 1),
            bounce_rate = single_page / n() * 100) -> c_bounce_rate
  View()
  
write.csv(c_bounce_rate, file = "F:/JAVERIA FIAZ/June 2023/Portfolio Project 2/csv/c_bounce_rate.csv", row.names = FALSE)

```

Show Traffic by each campaign
```{r for haet map display}
df_website_sessions  |> 
  mutate(hour = hour(created_at),
         hour_12 = format(created_at, "%I %p")) |> 
  group_by(utm_campaign, hour_12)  |> 
  summarize(count = n()) -> c_traffic_time
  
  
write.csv(c_traffic_time, file = "F:/JAVERIA FIAZ/June 2023/Portfolio Project 2/csv/c_traffic_time.csv", row.names = FALSE)

```

## Content

Total users and total sessions by each content
```{r}
df_website_sessions |> 
  group_by(utm_content) |> 
  summarise(total_sessions = n_distinct(website_session_id),
            total_user = n_distinct(user_id)) -> utm_content
 view()

write.csv(utm_content, file = 
            "F:/JAVERIA FIAZ/June 2023/Portfolio Project 2/csv/utm_content.csv", row.names = FALSE)

```

Most used device_type by each content
````{r}
# Most used device_type by each content
df_website_sessions |> 
  group_by(utm_content, device_type) |>
  summarise(
   device_used = table(device_type)) -> con_device 
  view()
  
write.csv(con_device, file = 
            "F:/JAVERIA FIAZ/June 2023/Portfolio Project 2/csv/con_device.csv", row.names = FALSE)
  
```

Traffic to order conversion for utm_content
```{r}
con_traffic_order_cvr <- df_website_sessions %>%
  left_join(df_orders, by = "website_session_id") %>%
  group_by(utm_content) %>%
  summarize(
    sessions = n_distinct(website_session_id),
    orders = n_distinct(order_id),
    session_order_CVR = round((n_distinct(order_id) / n_distinct(website_session_id)*100), 1)
  ) %>%
  arrange(desc(session_order_CVR))

print(con_traffic_order_cvr)

write.csv(con_traffic_order_cvr, file = "F:/JAVERIA FIAZ/June 2023/Portfolio Project 2/csv/con_traffic_order_cvr.csv", row.names = FALSE)

```

Show Traffic by each content
```{r for haet map display}
df_website_sessions  |> 
  mutate(hour = hour(created_at),
         hour_12 = format(created_at, "%I %p")) |> 
  group_by(utm_content, hour_12)  |> 
  summarize(count = n()) -> con_traffic_time
  
  
write.csv(con_traffic_time, file = "F:/JAVERIA FIAZ/June 2023/Portfolio Project 2/csv/con_traffic_time.csv", row.names = FALSE)

```

## Categorize: use of devices by utm_content per quarter of year
```{r}
df_website_sessions |> 
  mutate(yr = year(created_at),
         mon = months(created_at)) |> 
  group_by(yr, mon, utm_content, device_type) |> 
  summarise(
   device_used = table(device_type)) -> con_device_yr_mon
  view()
  

write.csv(con_device_yr_mon, file = "F:/JAVERIA FIAZ/June 2023/Portfolio Project 2/csv/con_device_yr_mon.csv", row.names = FALSE)

```


## Analyzing quarterly session to revenue per order 
```{r}

df_orders |> 
  mutate(yr = year(created_at),
         qtr = quarter(created_at)) |> 
  group_by(yr, qtr)  |> 
  summarize(revenue_per_order = round(sum(price_usd) / n_distinct(df_orders$order_id), 2),
    revenue_per_user =  round(sum(price_usd)/ n_distinct(user_id), 2)
    ) -> revenue_per_ord_usr

write.csv(revenue_per_ord_usr, file = "F:/JAVERIA FIAZ/June 2023/Portfolio Project 2/csv/revenue_per_ord_usr.csv", row.names = FALSE)

```

Order by time
```{r}
df_orders |> 
  mutate(hour_12 = format(created_at, "%I %p"),
         created_at = as.Date(created_at),
         yr = year(created_at),
         mon = months(created_at), 
         date = day(created_at),
         day_in_week = weekdays(created_at)
         ) |> 
  group_by(date, hour_12, day_in_week, mon, yr) |> 
  summarize(count = n()) -> order_by_time
  view()
  
  write.csv(order_by_time, file = "F:/JAVERIA FIAZ/June 2023/Portfolio Project 2/csv/order_by_time.csv", row.names = FALSE)
  
  
  
  df_orders |> 
  mutate(hour_12 = format(created_at, "%I %p"),
         created_at = as.Date(created_at),
         yr = year(created_at),
         mon = months(created_at), 
         date = day(created_at),
         day_in_week = weekdays(created_at)
         ) |> 
  group_by(date, mon) |> 
  summarize(count = n()) |> 
    view()
  
    df_orders |> 
  mutate(hour_12 = format(created_at, "%I %p"),
         created_at = as.Date(created_at),
         yr = year(created_at),
         mon = months(created_at), 
         date = day(created_at),
         day_in_week = weekdays(created_at)
         ) |> 
  group_by(hour_12, day_in_week) |> 
  summarize(count = n()) |> 
    view()
  
         
```


## Revenue by product and year
```{r}

revenue_by_product_ <- df_order_item %>%
  mutate(created_at = as.Date(created_at)) %>%
  group_by(yr = year(created_at), mo = month(created_at)) %>%
  summarize(mrfuzzy_rev = sum(ifelse(product_id == 1, price_usd, 0)),
            lovebear_rev = sum(ifelse(product_id == 2, price_usd, 0)),
            birthdaybear_rev = sum(ifelse(product_id == 3, price_usd, 0)),
            minibear_rev = sum(ifelse(product_id == 4, price_usd, 0)),
            total_revenue = sum(price_usd),
            total_margin = sum(price_usd - cogs_usd)) %>%
  arrange(yr, mo)

 write.csv(revenue_by_product_, file = "F:/JAVERIA FIAZ/June 2023/Portfolio Project 2/csv/revenue_by_product.csv", row.names = FALSE)

```

```{r}

df_order_item %>%
  mutate( created_at = as.Date(created_at),
         yr = year(created_at),
         mon = months(created_at), 
    product_name = case_when(
      product_id == 1 ~ "mrfuzzy",
      product_id == 2 ~ "lovebear",
      product_id == 3 ~ "birthdaybear",
      product_id == 4 ~ "minibear",
    )
  )|> 
  group_by(product_name, yr,mon) |> 
  summarise(expense = sum(cogs_usd),
            income = sum(price_usd),
            gain = sum(price_usd - cogs_usd) ) -> by_product

 write.csv(by_product, file = "F:/JAVERIA FIAZ/June 2023/Portfolio Project 2/csv/by_product.csv", row.names = FALSE)


```
